
This work presents an R-based workflow for assessing the impact of different Digital Elevation Models (DEMs) on hydrological model performance across multiple watersheds in Ethiopia, Estonia, Spain, and the United States. Simulated monthly streamflow from several DEM products is evaluated against observed discharge using boxplots, hydrographs, and performance metrics including normalized RMSE and the index of agreement.

Part 0: Load Libraries and Input Data
```{r, message=FALSE, warning=FALSE}
# Load required libraries
library(readxl)
library(tidyverse)
library(ggplot2)
library(extrafont)
library(hydroGOF)
library(ie2misc)

# Define file path
data_path <- "...data/discharge.xlsx"

# Import basin datasets
basins <- list(
  ESP = read_excel(data_path, sheet = "esp"),
  EST = read_excel(data_path, sheet = "est"),
  ETH = read_excel(data_path, sheet = "eth"),
  USA = read_excel(data_path, sheet = "usa")
)

# DEM order (used consistently throughout)
dem_levels <- c(
  "Observed", "AW3D30", "COP30",
  "HydroSHEDS", "MERIT", "NASADEM", "TanDEM-X"
)

# Color palette
dem_colors <- c(
  '#1f77b4', '#ff7f0e', '#2ca02c',
  '#d62728', '#9467bd', '#8c564b', '#e377c2'
)
```

Part 1: Monthly Discharge Distribution (Boxplots)
```{r, message=FALSE, warning=FALSE}
plot_boxplot <- function(df, basin_name) {

  df_long <- df %>%
    pivot_longer(
      cols = Observed:`TanDEM-X`,
      names_to = "Product",
      values_to = "Value"
    ) %>%
    mutate(Product = factor(Product, levels = dem_levels))

  p <- ggplot(df_long, aes(x = Product, y = Value, fill = Product)) +
    geom_boxplot(width = 0.25) +
    stat_boxplot(geom = "errorbar") +
    ylab(bquote("Discharge"~(m^3~s^-1))) +
    xlab("") +
    theme_grey(base_size = 19) +
    theme(legend.position = "none") +
    scale_fill_manual(values = dem_colors)

  ggsave(
    filename = paste0("C:/dev/papers/output/dem/boxplot_", basin_name, ".png"),
    plot = p,
    width = 11, height = 4, dpi = 300
  )
}

# Generate boxplots for all basins
lapply(names(basins), function(b) plot_boxplot(basins[[b]], b))
```

Part 2: Temporal Discharge Comparison (Hydrographs)
```{r}
plot_hydrograph <- function(df, basin_name, show_legend = FALSE) {

  df_long <- df %>%
    pivot_longer(
      cols = Observed:`TanDEM-X`,
      names_to = "Product",
      values_to = "Value"
    ) %>%
    mutate(
      Product = factor(Product, levels = dem_levels),
      Date = as.Date(Date)
    )

  p <- ggplot(df_long, aes(Date, Value, color = Product)) +
    geom_line() +
    ylab(bquote("Discharge"~(m^3~s^-1))) +
    xlab("Time (year-month)") +
    theme_grey(base_size = 19) +
    scale_x_date(date_breaks = "2 years", date_labels = "%Y-%m") +
    scale_colour_manual(values = dem_colors)

  if (!show_legend) {
    p <- p + theme(legend.position = "none")
  } else {
    p <- p +
      guides(colour = guide_legend(ncol = 7)) +
      theme(legend.position = "bottom")
  }

  ggsave(
    filename = paste0("C:/dev/papers/output/dem/discharge_", basin_name, ".png"),
    plot = p,
    width = 11, height = 4, dpi = 300
  )
}
```

Part 3: Performance Metrics — Normalized RMSE (nRMSE)
```{r}
compute_nrmse <- function(df) {

  obs <- df$Observed
  obs_range <- range(obs)

  dem_cols <- setdiff(names(df), c("Date", "Observed"))

  nrmse <- sapply(dem_cols, function(dem) {
    rmse_val <- rmse(obs, df[[dem]])
    100 * (rmse_val / (obs_range[2] - obs_range[1]))
  })

  return(round(nrmse, 2))
}

# Compute nRMSE for each basin
nRMSE_results <- lapply(basins, compute_nrmse)

# View results
nRMSE_results
```

Part 4: Index of Agreement (d-index)
```{r}
compute_d_index <- function(df) {

  obs <- df$Observed
  dem_cols <- setdiff(names(df), c("Date", "Observed"))

  d_index <- sapply(dem_cols, function(dem) {
    dr(obs, df[[dem]])
  })

  return(round(d_index, 3))
}

# Compute d-index for each basin
d_index_results <- lapply(basins, compute_d_index)

# View results
d_index_results
```

##### END  ####




